---
title: "Worksheet #2"
author: "Kenan Jake Jimenez"
date: "2025-09-29"
output: pdf_document
---
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Worksheet in R Programming

---

## Exercise 1. Create a vector using `:` operator

**a. Sequence from -5 to 5**

```{r}
vec1 <- -5:5
vec1
```

Output:
[1] -5 -4 -3 -2 -1 0 1 2 3 4 5

*The output shows a sequence from -5 up to 5.*

**b. `x <- 1:7`**

```{r}
x <- 1:7
x
```

Output:
[1] 1 2 3 4 5 6 7

---

## Exercise 2. Create a vector using `seq()` function

**a. `seq(1, 3, by=0.2)`**

```{r}
seq_vec <- seq(1, 3, by = 0.2)
seq_vec
```

Output:
[1] 1.0 1.2 1.4 1.6 1.8 2.0 2.2 2.4 2.6 2.8 3.0

*The numbers increase from 1 to 3 by increments of 0.2.*

---

## Exercise 3. Census of Workers

```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
          22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43,
          53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48,
          27, 39, 19, 30, 61, 54, 58, 26, 18)
```

**a. Access 3rd element**

```{r}
ages[3]
```

Output: [1] 22

**b. Access 2nd and 4th elements**

```{r}
ages[c(2,4)]
```

Output: [1] 28 36

**c. Exclude 1st element**

```{r}
ages[2:50]
```

---

## Exercise 4. Named Vector

```{r}
vec_named <- c("first"=3, "second"=0, "third"=9)
vec_named
```

**a. Access first and third**

```{r}
vec_named[c("first","third")]
```

Output:
first third
3     9

---

## Exercise 5. Modify Sequence

```{r}
seq2 <- -3:2
seq2
```

Output: [1] -3 -2 -1 0 1 2

**a. Modify 2nd element → 0**

```{r}
seq2[2] <- 0
seq2
```

Output: [1] -3 0 -1 0 1 2

---

## Exercise 6. Diesel Fuel Purchased by Mr. Cruz

```{r}
month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun")
price <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
liters <- c(25, 30, 40, 50, 10, 45)

fuel_data <- data.frame(month, price, liters)
fuel_data
```

**a. Data Frame Output**

| Month | Price | Liters |
| ----- | ----- | ------ |
| Jan   | 52.50 | 25     |
| Feb   | 57.25 | 30     |
| Mar   | 60.00 | 40     |
| Apr   | 65.00 | 50     |
| May   | 74.25 | 10     |
| Jun   | 54.00 | 45     |

**b. Average Fuel Expenditure**

```{r}
avg_exp <- weighted.mean(price, liters)
avg_exp
```

Output: [1] 59.2625

---

## Exercise 7. Built-in Dataset: `rivers`

```{r}
river_stats <- c(
  length = length(rivers),
  sum = sum(rivers),
  mean = mean(rivers),
  median = median(rivers),
  variance = var(rivers),
  sd = sd(rivers),
  min = min(rivers),
  max = max(rivers)
)
river_stats
```

Output:
length   sum   mean   median   variance   sd   min   max
141   591952   591.2   425   98382.96   313.83   135   3710

---

## Exercise 8. Forbes Top Celebrities

```{r}
power_rank <- 1:25
celebrity <- c(
  "Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
  "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
  "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
  "Paul McCartney", "George Lucas", "Elton John", "David Letterman",
  "Phil Mickelson", "J.K. Rowling", "Brad Pitt", "Peter Jackson",
  "Dr. Phil McGraw", "Jay Leno", "Celine Dion", "Kobe Bryant"
)

pay <- c(
  67, 90, 225, 110, 90,
  332, 302, 41, 52, 88,
  55, 44, 55, 40, 233,
  34, 40, 47, 75, 25,
  39, 45, 32, 40, 31
)

forbes <- data.frame(power_rank, celebrity, pay)
forbes
```

**b. Update J.K. Rowling**

```{r}
jk_index <- which(forbes$celebrity == "J.K. Rowling")
forbes$power_rank[jk_index] <- 15
forbes$pay[jk_index] <- 90
forbes[jk_index, ]
```

Output:
power_rank celebrity       pay
15         J.K. Rowling   90

**c. Interpretation**
The output shows celebrities ranked by power and pay. After updating, J.K. Rowling is ranked 15th with a pay of 90 million dollars, putting her among other top earners. This shows how changes in pay can affect overall ranking.

---
