---
title: 'Work sheet #3b'
author: "Kenan Jake Jimenez"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Create a data frame using the table below.

a. Write the codes.
```{r}
household_df <- data.frame(
  sex = c("Male", "Female", "Female", "Male", "Male",
          "Female", "Female", "Male", "Female", "Male",
          "Male", "Female", "Male", "Female", "Male",
          "Female", "Male", "Female", "Female", "Male"),
  
  fathers_occupation = c("Farmer", "Driver", "Others", "Farmer", "Driver",
                        "Farmer", "Others", "Driver", "Farmer", "Others",
                        "Driver", "Farmer", "Driver", "Others", "Farmer",
                        "Driver", "Others", "Farmer", "Driver", "Others"),
  
  persons_at_home = c(5, 6, 4, 7, 5,
                      8, 3, 6, 5, 4,
                      6, 5, 4, 7, 5,
                      6, 5, 4, 3, 4),
  
  siblings_at_school = c(3, 5, 2, 6, 4,
                     7, 3, 5, 4, 2,
                     6, 5, 3, 7, 4,
                     6, 5, 4, 2, 3),
    
  type_of_house = c("Wood", "Concrete", "Semi-Concrete", "Wood", "Concrete",
                    "Wood", "Semi-Concrete", "Concrete", "Wood", "Semi-Concrete",
                    "Concrete", "Wood", "Semi-Concrete", "Concrete", "Wood",
                    "Semi-Concrete", "Concrete", "Wood", "Concrete", "Semi-Concrete")

)

household_df

```

b. Describe the data. Get the structure or the summary of the data
```{r}
str(household_df)
summary(household_df)
```

c. Is the mean number of siblings attending is 5?
```{r}
mean(household_df$siblings_at_school)
```

d. Extract the 1st two rows and then all the columns using the subsetting functions. Write the codes and its output.
```{r}
household_df[1:2, ]
```

e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
household_df[c(3,5), c(2,4)]
```

f. Select the variable types of houses then store the vector that results as types_houses.
Write the codes.
```{r}
type_houses <- household_df$type_of_house
type_houses
```
g. Select only all Males respondent that their father occupation was farmer. Write the
codes and its output.
```{r}
subset_males_farmer <- subset(household_df, sex == "Male" & fathers_occupation == "Farmer")
subset_males_farmer
```

h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. Write the codes and its outputs.
```{r}
subset(household_df, sex == "Female" & siblings_at_school >= 5)
```

#2. Write a R program to create an empty data frame. Using the following codes:
a. Describe the results.
```{r}
df <- data.frame(
Ints = integer(),
Doubles = double(),
Characters = character(),
Logicals = logical(),
Factors = factor(),
stringsAsFactors = FALSE
)

str(df)
```

#3. Create a .csv file of this. Save it as HouseholdData.csv

a. Import the csv file into the R environment. Write the codes.
```{r}
write.csv(household_df, "HouseholdData.csv", row.names = FALSE)

data <- read.csv("HouseholdData.csv")
data
```

b. Convert the Sex into factor using factor() function and change it into integer.[Legend: Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
data$sex <- factor(data$sex, levels = c("Male", "Female"), labels = c(1, 2))
data$sex <- as.integer(data$sex)
data
```

c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood
= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r} 
data$type_of_house <- factor(data$type_of_house,
levels = c("Wood", "Concrete", "Semi-Concrete"),
labels = c(1, 2, 3))
data$type_of_house <- as.integer(data$type_of_house)
data$type_of_house

```

d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What
is the R code and its output?
```{r}
data$fathers_occupation <- factor(data$fathers_occupation,
levels = c("Farmer","Driver","Others"),
labels = c(1,2,3))
data$fathers_occupation <- as.integer(data$fathers_occupation)
data
```

e. Select only all females respondent that has a father whose occupation is driver. Write
the codes and its output.
```{r}
subset(data, sex == 2 & fathers_occupation == 2)
```

f. Select the respondents that have greater than or equal to 5 number of siblings attending
school. Write the codes and its output.
```{r}
subset(data, siblings_at_school >= 5)
```

#4. Interpret the graph.
The graph in general indicates that the majority of respondents reside in wooden houses, then concrete and semi-concrete homes. Wooden house residents tend to have more school-going siblings, implying that big families are prevalent in such dwellings.