---
title: 'Work sheet #4b'
author: "Kenan Jake Jimenez"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1.The table below shows the data about shoe size and height. Create a data frame.
```{r}
shoe <- c(6.5,9.0,8.5,8.5,10.5,7.0,9.5,9.0,13.0,7.5,10.5,8.5,12.0,10.5,
          13.0,11.5,8.5,5.0,10.0,6.5,8.5,10.5,6.5,10.5,11.0,9.0,13.0)

height <- c(66.0,68.0,64.5,65.0,70.0,64.0,70.0,71.0,72.0,
            64.0,74.5,67.0,71.0,71.0,77.0,72.0,59.0,62.0,72.0,66.0,64.0,
            73.0,69.0,72.0,71.0,69.0,70.0)

gender <- c("F","F","F","M","M","F","F","F","F","M","M","F","M","M",
            "M","F","F","M","M","F","F","M","M","M","M","M","M")

df <- data.frame(shoe, height, gender)
df
```
#a. Describe the data.

#The dataset contains 27 respondents, and each person has three variables 
#recorded: shoe size, height, and gender. Shoe size and height are numerical 
#variables, while gender is a categorical factor with two levels 
#(Male and Female). The data reflects basic anthropometric measurements 
#commonly used to compare physical characteristics across groups.

#b. Create a subset by males and females with their corresponding shoe size and height.
#What its result? Show the R scripts.
```{r}
male_data <- subset(df, gender == "M", select = c(shoe, height, gender))
female_data <- subset(df, gender == "F", select = c(shoe, height, gender))

male_data
female_data
```
#c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
#result.
```{r}
mean_shoe <- mean(df$Shoe_size)
mean_height <- mean(df$Height)

mean_shoe
mean_height
```
#d. Is there a relationship between shoe size and height? Why?

Yes. In general, larger shoe sizes tend to appear with greater heights in the 
dataset. This relationship is expected because shoe size is partly correlated
with body proportions, and taller individuals typically have larger feet. 
While this dataset shows a pattern, a statistical method such as correlation analysis #would confirm the strength of the relationship more rigorously.

#2. Construct character vector months to a factor with factor() and assign the result to
#factor_months_vector. Print out factor_months_vector and assert that R prints out
#the factor levels below the actual values.
```{r}
months_vector <- c(
  "March","April","January","November","January",
  "September","October","September","November","August",
  "January","November","November","February","May","August",
  "July","December","August","August","September","November","February",
  "April"
)
factor_months_vector <- factor(months_vector)
factor_months_vector
```

#3. Then check the summary() of the months_vector and factor_months_vector. | Inter-
#pret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
summary(factor_months_vector)
```
#No. The character vector simply stores text and cannot summarize categories.
#The factor version groups identical values and provides counts, which is more useful for analysis especially for categorical data like months.
#Therefore, the factor is much more useful when studying frequency, patterns, or performing statistical analysis.

#4. Create a vector and factor for the table below.
```{r}
factor_data <- c("East", rep("West", 4), rep("North", 3))
factor_data

new_order_data <- factor(factor_data, levels = c("East", "West", "North"))
print(new_order_data)

```

#5. Enter the data below in Excel with file name = import_march.csv
```{r}
import_march <- read.table("import_march.csv", header = TRUE, sep = ",")
import_march

```

#Using Conditional Statements (IF-ELSE)
#6. Full Search
```{r}
num <- as.integer(readline("Select a number from 1 to 50: "))

cat("Chosen number:", num, "\n")

if (is.na(num)) {
  cat("Invalid input. Please enter a number.\n")
  
} else if (num < 1 || num > 50) {
  cat("The number selected is beyond the range of 1 to 50\n")
  
} else if (num == 20) {
  print(TRUE)
  
} else {
  print(num)
}

```

#7. Change
```{r}
min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  
  remaining <- price
  count <- 0
  
  for (b in bills) {
    if (remaining >= b) {
      count <- count + (remaining %/% b)
      remaining <- remaining %% b
    }
  }
  
  return(count)
}

```

#8. The following is each student’s math score for one semester. Based on this, answer the
#following questions.

#a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
students <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

students
```
#b.
```{r}
for (i in 1:nrow(students)) {
  avg <- (students$Grade1[i] + students$Grade2[i] +
          students$Grade3[i] + students$Grade4[i]) / 4
  
  if (avg > 90) {
    cat(students$Name[i], "'s average grade this semester is ", avg, ".\n", sep="")
  }
}
```

#c.
```{r}
tests <- students[, 2:5]

for (j in 1:4) {
  total <- sum(tests[, j])
  avg <- total / nrow(tests)

  if (avg < 80) {
    cat("The", j, "th test was difficult.\n")
  }
}
```
#d
```{r}
for (i in 1:nrow(students)) {
  grades <- students[i, 2:5]
  highest <- grades[1]

  for (g in grades) {
    if (g > highest) highest <- g
  }
  
  if (highest > 90) {
    cat(students$Name[i], "'s highest grade this semester is ", highest, ".\n", sep="")
  }
}

```