---
title: 'Work sheet #4b'
author: "Kenan Jake Jimenez"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1.The table below shows the data about shoe size and height. Create a data frame.
```{r}
S <- c(6.5,9.0,8.5,8.5,10.5,7.0,9.5,9.0,13.0,7.5,10.5,8.5,12.0,10.5,
          13.0,11.5,8.5,5.0,10.0,6.5,8.5,10.5,6.5,10.5,11.0,9.0,13.0)

H <- c(66.0,68.0,64.5,65.0,70.0,64.0,70.0,71.0,72.0,
            64.0,74.5,67.0,71.0,71.0,77.0,72.0,59.0,62.0,72.0,66.0,64.0,
            73.0,69.0,72.0,71.0,69.0,70.0)

G <- c("F","F","F","M","M","F","F","F","F","M","M","F","M","M",
            "M","F","F","M","M","F","F","M","M","M","M","M","M")

DF <- data.frame(S, H, G)
DF
```

#a. Describe the data.
```{r}
"The dataset is made up of 27 entries, and for each person, three types of information were collected: how tall they are, their shoe measurement, and whether they are male or female. Height and shoe size are both recorded as numeric values, while gender is listed as a category with two possible labels. Overall, it’s a simple set of physical measurements that can be used to compare traits between the two groups."
```

#b. Create a subset by males and females with their corresponding shoe size and height.
#What its result? Show the R scripts.
```{r}
M_data <- subset(DF, G == "M", select = c(S, H, G))
F_data <- subset(DF, G == "F", select = c(S, H, G))

M_data
F_data
```
#c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
#result.
```{r}
S_mean <- mean(DF$S)
H_mean <- mean(DF$H)

S_mean
H_mean
```

#d. Is there a relationship between shoe size and height? Why?
```{r}
"Yeah, in general, people with bigger shoe sizes in the dataset also tend to be taller. That makes sense because shoe size is usually related to body proportions—taller people often have bigger feet. You can see the pattern in the data, but if you want to be sure how strong the relationship really is, you’d need to run something like a correlation analysis."
```

#2. Construct character vector months to a factor with factor() and assign the result to
#factor_months_vector. Print out factor_months_vector and assert that R prints out
#the factor levels below the actual values.
```{r}
M_vector <- c(
  "March","April","January","November","January",
  "September","October","September","November","August",
  "January","November","November","February","May","August",
  "July","December","August","August","September","November","February",
  "April"
)
factor_Mvector <- factor(M_vector)
factor_Mvector
```

#3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(M_vector)
summary(factor_Mvector)

"No, A regular character vector is basically just a list of words, it doesn’t tell you how often something shows up. But once you turn it into a factor, R automatically groups the repeated values and keeps track of their counts. That makes it way more useful when you’re dealing with categories, like months, and want to check patterns or how many times each one appears. So overall, the factor format is a lot more helpful for actual analysis."
```

#4. Create a vector and factor for the table below.
```{r}
F_data <- c("East", rep("West", 4), rep("North", 3))
F_data

New_data <- factor(F_data, levels = c("East", "West", "North"))
print(New_data)

```

#5. Enter the data below in Excel with file name = import_march.csv
```{r}
import_march <- read.table("import_march.csv", header = TRUE, sep = ",")
import_march

```

#Using Conditional Statements (IF-ELSE)
#6. Full Search
```{r}
num <- as.integer(readline("Select a number from 1 to 50: "))

cat("Chosen number:", num, "\n")

if (is.na(num)) {
  cat("Invalid input. Please enter a number.\n")
  
} else if (num < 1 || num > 50) {
  cat("The number selected is beyond the range of 1 to 50\n")
  
} else if (num == 20) {
  print(TRUE)
  
} else {
  print(num)
}

```

#7. Change
```{r}
Min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  
  remaining <- price
  count <- 0
  
  for (b in bills) {
    if (remaining >= b) {
      count <- count + (remaining %/% b)
      remaining <- remaining %% b
    }
  }
  
  return(count)
}

```

#8. The following is each student’s math score for one semester. Based on this, answer the following questions.

#a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
students <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

students
```
#b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output.
```{r}
for (i in 1:nrow(students)) {
  avg <- (students$Grade1[i] + students$Grade2[i] +
          students$Grade3[i] + students$Grade4[i]) / 4
  
  if (avg > 90) {
    cat(students$Name[i], "'s average grade this semester is ", avg, ".\n", sep="")
  }
}
```

#c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests.
```{r}
tests <- students[, 2:5]

for (j in 1:4) {
  total <- sum(tests[, j])
  avg <- total / nrow(tests)

  if (avg < 80) {
    cat("The", j, "th test was difficult.\n")
  }
}
```
#d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.
```{r}
for (i in 1:nrow(students)) {
  grades <- students[i, 2:5]
  highest <- grades[1]

  for (g in grades) {
    if (g > highest) highest <- g
  }
  
  if (highest > 90) {
    cat(students$Name[i], "'s highest grade this semester is ", highest, ".\n", sep="")
  }
}

```